---
title: "Assignment 2"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: zenburn
    df_print: tibble
  md_document:
    variant: gfm
---

# Introduction

The data used for this assignment is from a listing of **Toyota Corolla** offered for resale in the year **2004**, possibly in the **Netherlands**. The total number of entries in this dataset is ***1436***. 

Given that data is a subset of all cars in the Netherlands, it is expected that trends in this data are influenced by **Dutch policy**, therefore a discussion of such is of ***utmost importance*** so that we can narrow down the field of inquiry. Back then, ***tax rates were proportional to the weight of the vehicle***. So, weight and tax are expected to be **strongly correlated**. 

Additionally, diesel cars being heavier than petrol, were liable to **increased taxation**. However, since diesel fuel was ***cheaper enough*** that long-distance drivers still often chose diesel over petrol, which contributed to the demand for diesel vehicles, translating to the **resale value** of diesel cars. 

All in all, one can expect to see an **interplay** between Dutch policy that made diesel vehicles costly on one hand while the lower fuel cost made them more affordable. Both these ***opposing factors*** are expected to drive the resale prices of cars. 

Though it must be noted that there will always be **outliers** that dominate any dataset. For example:

• The **age** of a car  
• The **number of kilometers driven**  

These factors will eventually outweigh the fuel type.

For this particular assignment we will focus on:

• **Distribution of cars by fuel type** (categorical plot)  
• **Verification of the relationship between taxes and weight** (continuous plot) which we expect to be ***positively correlated***  

In the later part of the assignment other hypotheses will be explored to test the factors that drove the **resale value**.


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggExtra)
library(ggridges)
library(scales)
library(ggpubr)
library(scales)
library(ggtext) 
library(ggthemes)
```


```{r}
# load the data
dataPath = "./data/ToyotaCorolla.csv"
corolla <- read_csv(dataPath) |> janitor::clean_names()
head(corolla)
```

# Categorical Plots

## 1. Simple count plot of cars by fuel type

The fuel-type mix in the data shows ***significant imbalance***. Almost all the cars use **petrol** as fuel. **Diesel** is a distant second and **CNG** cars are rare. Apparently, Toyota offerings were mostly petrol or perhaps this may simply be an ***idiosyncrasy*** of the data, we shall perhaps never know. But the **bias** is worth noticing, since it means the rest of the plots will mostly reflect petrol cars unless we split or facet by fuel type. This context will be ***important*** while interpreting significant effects.



```{r}
corolla |> 
  ggplot(aes(fuel_type)) +
  geom_bar()
```


## 2. Pie chart as an alternative

To emphasize the ***extreme imbalance*** a pie chart can be used. Pie charts are usually **decried** over bar plots because humans are incapable of visually interpreting angles in the pie over the height of the bar. But here they will serve the purpose of ***exaggerating the imbalance*** in the data, namely the **dominance of petrol cars**.


```{r}
# Calculate percentages and label positions
fuel <- corolla |> 
  count(fuel_type) |> 
  mutate(
    pct = n / sum(n),
    label = paste0(fuel_type, "\n", round(pct * 100, 1), "%") # Create label with fuel type and percentage
  ) |> 
  arrange(desc(n)) |> 
  mutate(ypos = cumsum(pct) - 0.5 * pct) # Calculate cumulative positions for label placement
fuel
```


```{r}
fuel |>  
  ggplot(aes(x = "", y = pct, fill = fuel_type)) +
  geom_bar(stat = "identity") +
  coord_polar("y")
```



## 3. Enhanced donut chart with improved aesthetics

The pie chart by itself is not **aesthetically pleasing**. Not only that, there are a few things such as the **grid lines** and the **background** that can be cleaned up. As an improvement a ***donut chart*** can be used instead. As a good practice the following enhancements have been added:

• **Custom theme**  
• **Color palette**  
• **Annotations**  
• **Title and subtitle**  

```{r fig.height=4}
# Hole size
hsize <- 2

ggplot(fuel, aes(x = hsize, y = pct, fill = fuel_type)) +
  geom_col(color = "white") +
  geom_text(aes(label = label, y = ypos), x = hsize + 1, size = 4) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "GnBu") +
  xlim(c(0, hsize + 1)) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(size = 10, color = "grey30")
  ) +
  labs(
    title = "Toyota Fuel Type Distribution",
    subtitle = "Petrol dominates the market, with Diesel far behind and CNG almost absent"
  )
```


## Bonus : The same using ggpubr library

```{r}
library(ggpubr)

 
donut_plot <- ggdonutchart(fuel, "pct", label = "label", lab.font = c(4, "plain", "black"),
      fill = "fuel_type", color = "white", palette = "GnBu")
donut_plot
```


```{r fig.height=4}
donut_plot +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(size = 10, color = "grey30")
  ) +
  labs(
    title = "Toyota Fuel Type Distribution",
    subtitle = "Petrol dominates the market, with Diesel far behind and CNG almost absent\nn=1436"
  )
```


# Continuous Variables

## Trend between quarterly tax and weight

### 1. Scatter plot

The scatter plot reveals that ***tax increases as the weight goes up***, but the tax is not **proportional** to weight, rather there are ***tax slabs*** as we can see most of the points are clustered at **discrete values**. This suggests the next step to visualize this better: ***create tax buckets***.

```{r}
# relationship between weight and quarterly tax
corolla |> 
  ggplot(aes(quarterly_tax, weight)) +
  geom_point()
```


## 2. Boxplot of Tax Band vs Weight of car


```{r}
(corolla_subset <- corolla |>
  select(weight, quarterly_tax, fuel_type) |> # select the required columns
  mutate(
    tax_band = factor(case_when( # create tax bands 
    quarterly_tax < 100 ~ "low",
    between(quarterly_tax, 100, 200) ~ "medium",
    quarterly_tax >= 200 ~ "high"), levels = c("low", "medium", "high"), ordered = TRUE),
    weight_bin = factor(case_when(  # additionally create weight bins
      weight < 1150 ~ "light",
      weight >= 1150 ~ "heavy"),
      levels = c("light", "heavy"), ordered = TRUE)
    )
  )
```


```{r}
corolla_subset |> 
  ggplot(aes(tax_band, weight)) +
  geom_boxplot()
```



### 3. Refinement

The previous plot can be improved by adding:

• **Custom theme**  
• **Color palette**  
• **Annotations**  
• **Title and subtitle**  

Considering the ***imbalance in the data***, the **group size** has also been added to the plot so that viewers are ***cautious of interpreting the effects as significant*** without considering the group size.

```{r, fig.height=4}
corolla_subset |> 
  ggplot(aes(tax_band, weight, fill = tax_band)) +
  geom_boxplot(width = 0.5, linewidth = 0.3) +
  scale_fill_brewer(palette = "GnBu") +
  scale_x_discrete(labels = function(x){
    counts <- table(corolla_subset$tax_band)
    paste0(x, "\n(n=", counts[x], ")")
  } ) +
  labs(
    title = "Weight Distribution by Tax Band",
    subtitle = "Heavier vehicles pay more taxes. Though sample size may hinder interpretation",
    x = "Tax Band",
    y = "Weight (kg)"
  ) +
  theme_pubr() +  
  theme(
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y = element_text(margin = margin(r = 15)),
    plot.subtitle = element_text(margin = margin(b = 15), color = "grey30", size = 10),
    plot.title = element_text(size = 15, face = "bold")
  ) 
```




