---
title: "Assignment 2"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_float: true
    highlight: zenburn
    df_print: tibble
  md_document:
    variant: gfm
---

# Introduction

The data used for this assignment is from a listing of **Toyota Corolla** offered for resale in the year **2004**, possibly in the **Netherlands**. The total number of entries in this dataset is ***1436***. 

Given that data is a subset of all cars in the Netherlands, it is expected that trends in this data are influenced by **Dutch policy**, therefore a discussion of such is of ***utmost importance*** so that we can narrow down the field of inquiry. Back then, ***tax rates were proportional to the weight of the vehicle***. So, weight and tax are expected to be **strongly correlated**. 

Additionally, diesel cars being heavier than petrol, were liable to **increased taxation**. However, since diesel fuel was ***cheaper enough*** that long-distance drivers still often chose diesel over petrol, which contributed to the demand for diesel vehicles, translating to the **resale value** of diesel cars. 

All in all, one can expect to see an **interplay** between Dutch policy that made diesel vehicles costly on one hand while the lower fuel cost made them more affordable. Both these ***opposing factors*** are expected to drive the resale prices of cars. 

Though it must be noted that there will always be **outliers** that dominate any dataset. For example:

• The **age** of a car  
• The **number of kilometers driven**  

These factors will eventually outweigh the fuel type.

For this particular assignment we will focus on:

• **Distribution of cars by fuel type** (categorical plot)  
• **Verification of the relationship between taxes and weight** (continuous plot) which we expect to be ***positively correlated***  

In the later part of the assignment other hypotheses will be explored to test the factors that drove the **resale value**.


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggExtra)
library(ggridges)
library(scales)
library(ggpubr)
library(scales)
library(ggtext) 
library(ggthemes)
```


```{r}
# load the data
dataPath = "./data/ToyotaCorolla.csv"
corolla <- read_csv(dataPath) |> janitor::clean_names()
head(corolla)
```

# Categorical Plots

## 1. Simple count plot of cars by fuel type

The fuel-type mix in the data shows ***significant imbalance***. Almost all the cars use **petrol** as fuel. **Diesel** is a distant second and **CNG** cars are rare. Apparently, Toyota offerings were mostly petrol or perhaps this may simply be an ***idiosyncrasy*** of the data, we shall perhaps never know. But the **bias** is worth noticing, since it means the rest of the plots will mostly reflect petrol cars unless we split or facet by fuel type. This context will be ***important*** while interpreting significant effects.



```{r}
corolla |> 
  ggplot(aes(fuel_type)) +
  geom_bar()
```


## 2. Pie chart as an alternative

To emphasize the ***extreme imbalance*** a pie chart can be used. Pie charts are usually **decried** over bar plots because humans are incapable of visually interpreting angles in the pie over the height of the bar. But here they will serve the purpose of ***exaggerating the imbalance*** in the data, namely the **dominance of petrol cars**.


```{r}
# Calculate percentages and label positions
fuel <- corolla |> 
  count(fuel_type) |> 
  mutate(
    pct = n / sum(n),
    label = paste0(fuel_type, "\n", round(pct * 100, 1), "%") # Create label with fuel type and percentage
  ) |> 
  arrange(desc(n)) |> 
  mutate(ypos = cumsum(pct) - 0.5 * pct) # Calculate cumulative positions for label placement
fuel
```


```{r}
fuel |>  
  ggplot(aes(x = "", y = pct, fill = fuel_type)) +
  geom_bar(stat = "identity") +
  coord_polar("y")
```



## 3. Enhanced donut chart with improved aesthetics

The pie chart by itself is not **aesthetically pleasing**. Not only that, there are a few things such as the **grid lines** and the **background** that can be cleaned up. As an improvement a ***donut chart*** can be used instead. As a good practice the following enhancements have been added:

• **Custom theme**  
• **Color palette**  
• **Annotations**  
• **Title and subtitle**  

```{r fig.height=4}
# Hole size
hsize <- 2

ggplot(fuel, aes(x = hsize, y = pct, fill = fuel_type)) +
  geom_col(color = "white") +
  geom_text(aes(label = label, y = ypos), x = hsize + 1, size = 4) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "GnBu") +
  xlim(c(0, hsize + 1)) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(size = 10, color = "grey30")
  ) +
  labs(
    title = "Toyota Fuel Type Distribution",
    subtitle = "Petrol dominates the market, with Diesel far behind and CNG almost absent"
  )
```


## Bonus : The same using ggpubr library

```{r}
library(ggpubr)

 
donut_plot <- ggdonutchart(fuel, "pct", label = "label", lab.font = c(4, "plain", "black"),
      fill = "fuel_type", color = "white", palette = "GnBu")
donut_plot
```


```{r fig.height=4}
donut_plot +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(size = 10, color = "grey30")
  ) +
  labs(
    title = "Toyota Fuel Type Distribution",
    subtitle = "Petrol dominates the market, with Diesel far behind and CNG almost absent\nn=1436"
  )
```


# Continuous Variables

## Trend between quarterly tax and weight

### 1. Scatter plot

The scatter plot reveals that ***tax increases as the weight goes up***, but the tax is not **proportional** to weight, rather there are ***tax slabs*** as we can see most of the points are clustered at **discrete values**. This suggests the next step to visualize this better: ***create tax buckets***.

```{r}
# relationship between weight and quarterly tax
corolla |> 
  ggplot(aes(quarterly_tax, weight)) +
  geom_point()
```


## 2. Boxplot of Tax Band vs Weight of car


```{r}
(corolla_subset <- corolla |>
  select(weight, quarterly_tax, fuel_type) |> # select the required columns
  mutate(
    tax_band = factor(case_when( # create tax bands 
    quarterly_tax < 100 ~ "low",
    between(quarterly_tax, 100, 200) ~ "medium",
    quarterly_tax >= 200 ~ "high"), levels = c("low", "medium", "high"), ordered = TRUE),
    weight_bin = factor(case_when(  # additionally create weight bins
      weight < 1150 ~ "light",
      weight >= 1150 ~ "heavy"),
      levels = c("light", "heavy"), ordered = TRUE)
    )
  )
```


```{r}
corolla_subset |> 
  ggplot(aes(tax_band, weight)) +
  geom_boxplot()
```



### 3. Refinement

The previous plot can be improved by adding:

• **Custom theme**  
• **Color palette**  
• **Annotations**  
• **Title and subtitle**  

Considering the ***imbalance in the data***, the **group size** has also been added to the plot so that viewers are ***cautious of interpreting the effects as significant*** without considering the group size.

```{r, fig.height=4}
corolla_subset |> 
  ggplot(aes(tax_band, weight, fill = tax_band)) +
  geom_boxplot(width = 0.5, linewidth = 0.3) +
  scale_fill_brewer(palette = "GnBu") +
  scale_x_discrete(labels = function(x){
    counts <- table(corolla_subset$tax_band)
    paste0(x, "\n(n=", counts[x], ")")
  } ) +
  labs(
    title = "Weight Distribution by Tax Band",
    subtitle = "Heavier vehicles pay more taxes. Though sample size may hinder interpretation",
    x = "Tax Band",
    y = "Weight (kg)"
  ) +
  theme_pubr() +  
  theme(
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y = element_text(margin = margin(r = 15)),
    plot.subtitle = element_text(margin = margin(b = 15), color = "grey30", size = 10),
    plot.title = element_text(size = 15, face = "bold")
  ) 
```





```{r}
pal <- c(Petrol = "#0072B2", Diesel = "#009E73", CNG = "#E69F00")
corolla_subset |> 
  ggplot(aes(weight, tax_band, fill = tax_band)) +
  geom_boxplot() +
  facet_grid(fuel_type ~ .) 
```


```{r}
p <- ggboxplot(corolla_subset, x = "tax_band", y = "weight",
                color = "tax_band", palette = pal,
                add = "jitter", shape = "tax_band")
p
```



```{r}
corolla_subset |> 
  ggplot(aes(km, price)) +
  geom_point()
```

```{r}
corolla_subset |> 
  ggplot(aes(km, price, size = age_08_04, color = age_08_04)) +
  geom_point(alpha = 0.5)
```


```{r}
corolla_subset |> 
  ggplot(aes(km, price, color = fuel_type)) +
  geom_point(alpha = 0.5)
```


```{r}


ggMarginal(p, type = "density", fill = "transparent")
```



```{r}
corolla |> 
  ggplot(aes(km, age_08_04)) + 
  geom_density_2d(aes(color = after_stat(level)), size = 0.5) + 
  geom_point(aes(color = price), alpha = 0.5)

```



```{r}
corolla |>
  ggplot(aes(hp, cc)) + 
  geom_point()
```




```{r}


corolla_subset <- corolla_subset |> 
  mutate(group_var = paste(tax_band, weight_bin, sep = " - "))

# Basic ridge plot
ggplot(corolla_subset, aes(x = price, y = group_var)) +
  geom_density_ridges(alpha = 0.7, scale = 0.9)

```


```{r}
corolla_subset |> count(weight_bin)
```

```{r}
corolla |> count(automatic)
```

```{r}
group_cc_hp <- corolla |> 
  filter(cc < 16000) |> 
  group_by(cc, hp) |> 
  summarise(
    avg_weight = mean(weight, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    .groups = 'drop'
  )


ggplot(group_cc_hp, aes(x = cc, y = hp, size = avg_weight, color = median_price)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(2, 10), name = "Average Weight") +  # Adjust size range
  scale_color_gradient(low = "blue", high = "red", name = "Median Price")
```



```{r}
library(emojifont)
```

```{r}
search_emoji("car")
```

```{r}
ggplot() + geom_emoji("car", color='red') + theme_void()
```

```{r}
set.seed(2016-03-09)
fa <- fontawesome(c('fa-github', 'fa-weibo', 'fa-twitter', 'fa-android', 'fa-car-side'))
d <- data.frame(x=rnorm(20),
                y=rnorm(20),
                label=sample(fa, 20, replace=T))

ggplot(d, aes(x, y, color=label, label=label)) +
    geom_text(family='fontawesome-webfont', size=15)+
    xlab(NULL)+ylab(NULL) +
    theme(legend.text=element_text(family='fontawesome-webfont'))
```


```{r}
library(ggwaffle)
waffle_data <- waffle_iron(mpg, aes_d(group = class))

ggplot(waffle_data, aes(x, y, fill = group)) + 
  geom_waffle()
```









```{r}
library(dplyr)
library(ggpubr)
library(scales)
library(forcats)

fuel <- corolla |> 
  count(fuel_type) |> 
  mutate(pct = n / sum(n),
         label = paste0(fuel_type, " — ", percent(pct, accuracy = 1))) |> 
  arrange(desc(n)) %>%
  mutate(fuel_type = fct_inorder(fuel_type))

pal <- c(Petrol = "#0072B2", Diesel = "#009E73", CNG = "#E69F00")  # colorblind‑safe

p <- ggdonutchart(
  fuel, x = "n", label = "labs",
  fill = "fuel_type", color = "white",
  lab.pos = "out", lab.font = c(10, "bold", "black"), # outside labels for readability
  size = 2, radius = 1, label.size = 0.3
) +
  scale_fill_manual(values = pal) +
  labs(
    title = "Fuel type share — Toyota (NL, 2004)",
    subtitle = paste0("n = ", sum(fuel$n), " cars • showing percentages")
  ) +
  theme_void(base_size = 12) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )

p

```




```{r}
library(dplyr)
library(ggplot2)
library(scales)

fuel <- corolla %>%
  count(fuel_type) %>%
  mutate(pct = n/sum(n),
         label = paste0(fuel_type, " — ", percent(pct, 0.1)))

# Calculate cumulative positions for label placement
fuel <- fuel %>%
  arrange(desc(fuel_type)) %>%
  mutate(ypos = cumsum(pct) - 0.5*pct)

ggplot(fuel, aes(x = 2, y = pct, fill = fuel_type)) +
  geom_bar(stat = "identity", color = "white", width = 1) +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +  # adds space for outside labels
  geom_text(aes(y = ypos, label = label), color = "black", size = 4, x = 2.7) +
  scale_fill_manual(values = c("#0072B2", "#009E73", "#E69F00")) +
  theme_void() +
  theme(legend.position = "none") +
  labs(title = "Fuel type share — Toyota (NL, 2004)",
       subtitle = paste0("n = ", sum(fuel$n), " cars • showing percentages"))

```






```{r}
# packages
library(tidyverse)
library(ggtext)     # for emoji rendering in ggplot text
# Optional (if emoji don't render): showtext::showtext_auto(); sysfonts::font_add_google("Noto Color Emoji","noto")

# counts you gave
cnt <- c(Petrol = 1264, Diesel = 155, CNG = 17)
p   <- round(100 * cnt / sum(cnt))          # -> 88, 11, 1 (sums to 100)

# make 10x10 grid and assign fuel blocks
waffle <- tibble(
  fuel = rep(names(p), p),
  i = 0:99,
  x = i %% 10,
  y = 9 - i %/% 10
)

ggplot(waffle, aes(x, y, color = fuel)) +
  geom_text(label = "\U0001F697", size = 8) +  # 🚗
  scale_color_manual(values = c(Petrol = "#4C9AFF", Diesel = "#36B37E", CNG = "#FF5630")) +
  coord_equal(expand = FALSE) +
  labs(title = "Fuel type share (Toyota, NL 2004)",
       subtitle = sprintf("Petrol %d%%  •  Diesel %d%%  •  CNG %d%%", p["Petrol"], p["Diesel"], p["CNG"])) +
  theme_void(base_size = 12) +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold"))

```










































































































































































